### 2024-03-11 doc/15yrsChange.Rmd
Update Author List and Tweak Figure 4
Updated author list
Updated author list (Pt. 2)
cleaning R markdown for export to google doc & figure 5 tweaks

**Authors (tentative / incomplete list, not in order):**

@@

*corresponding author: MorrealeLL@si.edu; +1 607 227 1350

@@

#Figures
map_and_photos = 1
photo_deer_exclosure = "1a"
photo_represenative_upland = "1b"
photo_graveyard = "1c"

@@

mortality_by_genera = 4 # 4-panel figure showing biomass mortality by genera for for (a) whole plot, (b) deer exclosure, (c) upland, (d) graveyard

regen_failure = 5 #some figure illustrating absence of small canopy trees/ shifting  size distributions

@@

The temperate forest biome plays a critical role in the global carbon cycle, accounting for almost half of the net global forest carbon (C) sink [@harris_global_2021], with deciduous forests representing a substantial portion of this, sequestering >300 Tg C yr^-1^ [@pugh_role_2019]. 
Although currently a C sink, the future of the biome remains uncertain.
The dominant view imbued in Earth System Models is that, at least within the eastern US, the biome is likely to remain a C sink for the remainder of the 21st century [@finzi_carbon_2020; @wu_uncertainty_2023], albeit with declining CO~2~ sequestration capacity [@ahlström_robustness_2012]. 
Yet, global C models predict a wide range of future trajectories of CO~2~ sequestration [@ahlström_robustness_2012;@arora_carbon_2020]<!--check refs-->, and current global C models do not represent some influential mechanisms [e.g., @fatichi_moving_2014; @clark_continentwide_2021].
One mechanism that is not represented in global C models is the impact of nuisance species -- i.e., endemic or non-endemic species that, as a result of human influence, have much greater abundance in an ecosystem than they did historically, resulting in undesirable ecological consequences [@refs]. 
Similarly, the impact of nuisance species on forest carbon budgets is not considered in machine learning/niche models that seek to project future forest distribution and carbon stocks [**?** @wu_uncertainty_2023], nor in carbon offset projects [**???** @refs].
This is problematic in that nuisance species -- including non-endemic insect pests and pathogens, non-endemic plants, and over-abundant herbivores -- are dramatically impacting carbon cycling in temperate deciduous forests around the world, as described below.
"Tree mortality is one of the largest uncertainities in projections of future forest dynamics" (Bugmann et al. 2019) - (*Add reference*)

## (Canopy trees and nonindigenous pests and pathogens)

Nuisance pests and pathogens have had a protracted and enduring impact on forests, often modifying forest productivity, nutrient cycling, successional trajectories and geomorphic processes (Lovett et al. 2016, Fischer et al. 2013, Herms and McCullough 2014, Fei et al. 2014).  The presence of nonindigenous forest insects and diseases, which have been historically introduced to new ecosystems through intercontinental trade, will continue to proliferate with increased globalization and human movement (Fischer et al. 2013, Aukema et al. 2010, Brockerhoff et al. 2006). Many of the prominent pests and pathogens in temperate deciduous forests have devastating effects on canopy speciation and structure.  Once abundant in temperate deciduous forests throughout North America, the American Chestnut (*Castanea dentata*) and the Ash genera (*Fraxinus spp.*) have been decimated by pests and pathogens, with stand-clearing disturbances occuring within a few years after initial infestation (Anderson 2020, Klooster et al. 2013, Need a source for Ash abundance and length of time between infestation and mortality for Chestnuts). These previously dominant overstorey species are now confined to the understory as non-reproductive seeds and saplings (Elliot and Swank 2008, Anagnostakis 1987, Need a source for Ash).  In other species, such as the Eastern Hemlock (*Tsuga canadensis*), the existence of pests and pathogens induces close to total population mortality, with considerable risk of regional extirpation (Ellison et al. 2018).  The loss of mature canopy species to nuisance species contributes to global change driven biomass losses, with biomass losses due to pest and pathogens comparable to those attributed to fire and forest removal (Fei et al. 2013)  Since many pests and pathogens do not currently occupy the full extent of their host species range, a significant portion of the remaining forest canopy is susceptible to future invasion from nuisance species, further increasing potentials for biomass loss and altered carbon cycle dynamics (Fei et al. 2013)

@@

## (paragraph on recruitment failure because of deer and non-endemic plants)

Recruitment Failure Paragraph

@@

Here, we use 15 years of detailed census data from a large forest dynamics plot including a 4-ha deer exclosure to test the following hypotheses:
(1) C stocks have declined (-∆AGB)
(2) canopy tree mortality and associated biomass loss are increasing, in large part due to non-endemic nuisance species (pests & pathogens)
(3) recruitment (outside deer exclosure) and growth have not kept pace with tree mortality, implying that future mortality will result in substantial net biomass loss (regeneration debt)

@@

*Need to add the Anderson-Teixeira et al. 2015 reference to references page*

The study was conducted in the 25.6-hectare Large Forest Dynamics plot at the National Zoo and Conservation Biology Institute in Front Royal, Virginia (38 53’36.6“N, 78 08’43.4 “W).  The plot, which is located in the central Appalachian Mountains adjacent to Shenandoah National Park, is composed of mature secondary eastern deciduous forest. Situated in the Appalachian Oak forest region, the canopy is dominated by tulip poplar (*Liriodendron tulipifera*), oak (*Quercus spp.*), and hickory (*Carya spp.*) and the understory is primarily composed of spicebush (*Lindera benzoin*), paw-paw (*Asimina triloba*), American hornbeam (*Carpinus caroliniana*) and witch hazel (*Hamamelis virginiana*) (SCBI site description). The land-use history of the site is varied, including periods of agricultural development and intensive logging, with dendrological data estimating canopy tree establishment around 1900 (SI archives, Bourg et al. 2013).  Natural disturbances at the plot consist of wind and infrequent ice storms (Anderson-Teixiera 2015).  The plot, which includes a 4-hectare deer exclosure that has decreased the presence of deer since 1990, is divided into 640 quadrats, each measuring 20 x 20 meters. It is one of 78 sites in the Forest Global Earth Observatory (ForestGEO), a global network of forest dynamic plots that promotes comparative forest ecology studies around the World.

*Add information about the history of pests and pathogens at this site and at similar forests*

In addition to our consideration of the plot as a whole, we give special consideration to three portions of the plot, all upland forest (i.e., excluding low-lying areas around streams):
(1) low deer, low non-endemic insects & pathogens (Fig. `r photo_deer_exclosure`)- 4 ha portion of upland forest in the SE quarter of the plot, fenced in YEAR and maintained with only occasional deer presence for the past ## years and with relatively low abundance of canopy species affected by non-endemic pests and pathogens;
(2) high deer, low non-endemic insects & pathogens (Fig. `r photo_represenative_upland`) - # ha of upland forest outside the deer exclosure with relatively low abundance of canopy species affected by non-endemic pests and pathogens; 
(3) high deer, high non-endemic insects & pathogens (Fig. `r photo_graveyard`) - a # ha section of upland forest outside the deer exclosure with relatively high abundance of canopy species affected by non-endemic pests and pathogens. 
These areas were delineated as follows.
Upland forest was defined according to topographic wetness index, originally calculated for the SCBI plot in @mcgregor_tree_2021.
Vulnerability to non-endemic insects & pathogens was defined based on the abundance of canopy species affected by non-endemic pests and pathogens at the time of plot establishment (2008).

![**Figure `r map_and_photos`. Photos within the SCBI ForestGEO plot: (a) low deer, low non-endemic insects & pathogens, (b) high deer, low non-endemic insects & pathogens, (c) high deer, high non-endemic insects & pathogens ** *(add a map of the plot)* All photos taken September 2023 by K. Anderson-Teixeira.](display/photos.png)

@@

*(Describe main census and mortality survey)*

A number of surveys are regularly conducted at the site, two of which are included in our analysis.  As part of the ForestGEO network, the plot undergoes a comprehensive woody plant inventory every five years, according to the protocol outlined in Condit (1998).  This inventory, hereby referred to as the census, includes all stems greater than 1 centimeter in diameter breast height (1.3 meters; hereafter referred to as dbh).  The census records information regarding the dbh, species, status and location of each stem included in the survey.  Each plant is assigned an identifying number for sequential data collection in subsequent censuses and outfitted with a metal tag in the field. For multi-stemmed individuals, each additional stem that surpasses the 1 cm dbh threshold receives a stem number and associated tag.  The location of each individual within its respective 20 x 20 meter quadrat is recorded on a map of the plot on ArcGIS FieldMaps.  Established in 2008, there have been 4 censuses at the site, comprising 20 years of detailed forest dynamics data. In addition to the ForestGEO census, an annual mortality survey is conducted on the site.   Through assessing individual tree health, this study illustrates trends in forest mortality and identifies factors that are associated with death.  Data is collected on the current status, canopy position, percentage of crown living and intact, and visible indicators of poor tree health, such as physical damage, potential pathogens and insect infestation.

*Add Condit 1998 in references*

*In response to the presence of Emerald Ashborer, a mortality survey was initiated, then improved in year [X].

*(describe seedlings survey, if used)*

*(describe QA/QC and technical workflow)*

@@

apply the metric of @miller_overabundant_2023 to determine imminent failure/ probable failure/ insecure/ secure?

@@

## Paragraph discussing broad trend of biomass stock decrease (Figure 2) & net change in biomass & differences among regions in the plot 
For the first time at the Front Royal ForestGEO site, whole-plot C stocks have decreased since the last census (Figure 2). From 2018 to 2023, whole-plot C stocks have decreased from **"X to X"** Mg C/ha. Both low-nuisance regions show similar decreases in C stocks going from **"X to X"** and **"X to X"** Mg C/ha, respectively. In the high deer high nuisance regions C stocks have decreased for the past two censuses, with a higher magnitude decrease of **"X to X"** Mg C/ha from 2018 to 2023. C stock decrease **is related to/ is explained by** above ground net biomass change. In the first two census intervals, 2008 to 2013 and 2013 to 2018, there were positive C fluxes, but for the 2018 to 2023 interval there was a negative C flux of **VALUE** Mg C/ha/year (Figure 3a).

## Paragraph discussing constituent fluxes of net biomass change - across the whole plot & within regions
The trend in net woody production can be explained by its constituent fluxes (Figure 3). Trends in woody mortality show that for the first two census intervals, the whole plot has lost **VALUE** and then **VALUE** Mg C/ha/year, increasing to a loss of **VALUE** Mg C/ha/year for the most recent census interval (Figure 3b). Woody productivity contributes positively to the flux, showing that the whole plot gained **VALUE** Mg C/ha/year in the first interval, **VALUE** Mg C/ha/year in the second, and **VALUE** Mg C/ha/year in the third (Figure 3c). Areas of low deer, low nuisance species were consistently the most productive ending at **VALUE** Mg C/ha/year in the interval 2018 to 2023, and areas of high deer, high nuisance species were the least productive at **VALUE** Mg C/ha/year in the interval 2018 to 2023. Recruitment has consistently added smaller magnitude amounts of carbon with the whole plot gaining **VALUE** Mg C/ha/year (Figure 3d). Recruitment was highest at **VALUE** Mg C/ha/year in areas with high deer and high vulnerable species, and lowest at **VALUE** Mg C/ha/year in areas with high deer and low vulnerable species.

## Species compostion of mortality flux

Woody mortality was the largest driver of the negative trend in biomass, and certain species were responsible for higher levels of mortality. Fraxinus americana had the highest aboveground woody mortality rates of any species present on the plot, peaking at **VALUE** Mg C/ha/year lost in *2020* , and **VALUE** Mg C/ha/year lost in 2022 (Figure 4). Quercus velutina contributed the second-highest levels of mortality, losing **VALUE** Mg C/ha/year in 2022. Quercus rubra, Carya species, Quercus prinus, Quercus alba, and Liriodendron tulipifera were all significant contributors as well.

## Species compositon of recruitment flux
Certain species dominated aboveground woody recruitment, with understory species being recruited at higher rates than canopy species (Figure 5). Out of the canopy species, Carya species showed the highest levels of recruitment with a rate of **VALUE** Mg C/ha/year recruited from 2018 to 2023 (Figure 5a). Acer negundo had the second highest rate of recruitment for canopy species at **VALUE** Mg C/ha/year from 2018 to 2023. Fagus grandifolia, Fraxinus americana, and Ulmus rubra are the other species with the highest recruitment rates. Most canopy species had lower recruitment rates in the 2018 to 2023 census interval than the 2008 to 2013 interval. Lindera benzoin had by far the highest recruitment rate of the understory species at **VALUE** Mg C/ha/year from 2018 to 2023 (Figure 3b). Asimina triloba had the second highest rate of recruitment of the understory species at **VALUE** Mg C/ha/year recruited from 2018 to 2023. Other understory species with high recruitment were Carpinus caroliniana, Hamamelis virginiana, and Celtis occidentalis.

@@

@@

# Conclusions (optional) *(No Conclusions section listed in Submission Guidelines)*

>>>>

**Authors :**

@@

*corresponding author: MorrealeLL@si.edu;

@@

#Figures
map_and_photos = 1

@@

mortality_by_genera = 4 # 2-panel figure showing annual biomass mortality by genera for (a) canopy species and (b) understory species

recruitment_by_genera = 5 # 2-panel figure showing # of recruited stems by genera for (a) canopy species and (b) understory species by plot regions

regen_failure = 6 #some figure illustrating absence of small canopy trees/ shifting  size distributions

@@

The temperate forest biome plays a critical role in the global carbon cycle, accounting for almost half of the net global forest carbon (C) sink [@harris_global_2021], with deciduous forests representing a substantial portion of this, sequestering >300 Tg C yr^-1^ [@pugh_role_2019]. 
Although currently a C sink, the future of the biome remains uncertain.
The dominant view imbued in Earth System Models is that, at least within the eastern US, the biome is likely to remain a C sink for the remainder of the 21st century [@finzi_carbon_2020; @wu_uncertainty_2023], albeit with declining CO~2~ sequestration capacity [@ahlström_robustness_2012]. 
Yet, global C models predict a wide range of future trajectories of CO~2~ sequestration [@ahlström_robustness_2012;@arora_carbon_2020]<!--check refs-->, and current global C models do not represent some influential mechanisms [e.g., @fatichi_moving_2014; @clark_continentwide_2021].
One mechanism that is not represented in global C models is the impact of nuisance species -- i.e., endemic or non-endemic species that, as a result of human influence, have much greater abundance in an ecosystem than they did historically, resulting in undesirable ecological consequences [@refs]. 
Similarly, the impact of nuisance species on forest carbon budgets is not considered in machine learning/niche models that seek to project future forest distribution and carbon stocks [**?** @wu_uncertainty_2023], nor in carbon offset projects [**???** @refs].
This is problematic in that nuisance species -- including nonindigenous insect pests and pathogens, nonindigenous plants, and over-abundant herbivores -- are dramatically impacting carbon cycling in temperate deciduous forests around the world, as described below.
"Tree mortality is one of the largest uncertainities in projections of future forest dynamics" (Bugmann et al. 2019)

Nuisance pests and pathogens can have a protracted and enduring impact on forests, often modifying forest productivity, nutrient cycling, successional trajectories and geomorphic processes (Lovett et al. 2016, Fischer et al. 2013, Herms and McCullough 2014, Fei et al. 2014).  The presence of nonindigenous forest insects and diseases, which have been historically introduced to new ecosystems through intercontinental trade, will continue to proliferate with increased globalization and human movement (Fischer et al. 2013, Aukema et al. 2010, Brockerhoff et al. 2006). Many of the prominent pests and pathogens in temperate deciduous forests have devastating effects on canopy composition and structure.  Once abundant in temperate deciduous forests throughout North America, the American Chestnut (*Castanea dentata*) and the Ash genera (*Fraxinus spp.*) have been decimated by pests and pathogens, with stand-clearing disturbances occuring within a few years after initial infestation (Anderson 2020, Klooster et al. 2013, Need a source for Ash abundance and length of time between infestation and mortality for Chestnuts). These previously dominant overstorey species are now confined to the understory as non-reproductive seeds and saplings (Elliot and Swank 2008, Anagnostakis 1987, Need a source for Ash).  In other species, such as the Eastern Hemlock (*Tsuga canadensis*), the existence of pests and pathogens induces close to total population mortality, with considerable risk of regional extirpation (Ellison et al. 2018).  The loss of mature canopy species to nuisance species contributes to global change driven biomass losses, with biomass losses due to pest and pathogens likely comparable to those attributed to fire and forest removal (Fei et al. 2013)  Since many pests and pathogens do not currently occupy the full extent of their host species range, a significant portion of the remaining forest canopy is susceptible to future invasion from nuisance species, further increasing potentials for biomass loss and altered carbon cycle dynamics (Fei et al. 2013)

@@

@@

Here, we use 15 years of detailed census data from a large forest dynamics plot including a 4-ha deer exclosure to test the following hypotheses:
(1) ) Aboveground C stocks in living biomass have declined since the initial census (-∆AGB)
(2) canopy tree mortality and associated woody biomass loss are increasing, in large part due to nonindigenous nuisance species
(3) Recruitment and woody growth have not kept pace with tree mortality, suggesting that future mortality will result in substantial net biomass loss and sustained decrease in the C storage potential of this ecosystem.

@@

The study was conducted in the 25.6-hectare Large Forest Dynamics plot at the National Zoo and Conservation Biology Institute in Front Royal, Virginia (38 53’36.6“N, 78 08’43.4 “W).  The plot, which is located in the central Appalachian Mountains adjacent to Shenandoah National Park, is composed of mature secondary eastern deciduous forest. Situated in the Appalachian Oak forest region, the canopy is dominated by tulip poplar (*Liriodendron tulipifera*), oak (*Quercus spp.*), and hickory (*Carya spp.*) and the understory is primarily composed of spicebush (*Lindera benzoin*), paw-paw (*Asimina triloba*), American hornbeam (*Carpinus caroliniana*) and witch hazel (*Hamamelis virginiana*) (SCBI site description). The land-use history of the site is varied, including periods of agricultural development and intensive logging, with dendrological data estimating canopy tree establishment around 1900 (SI archives, Bourg et al. 2013).  Natural disturbances at the plot consist of wind and infrequent ice storms (Anderson-Teixiera 2015).  The plot, which includes a four-hectare deer exclosure that has decreased the presence of deer since 1990, is divided into 640 quadrats, each measuring 20 x 20 meters. It is one of 78 sites in the Forest Global Earth Observatory (ForestGEO), a global network of forest dynamic plots that promotes comparative forest ecology studies around the world.

* Add information about the history of pests and pathogens at this site and at similar forests

In addition to our consideration of the plot as a whole, we give special consideration to three portions of the plot, all upland forest (i.e., excluding low-lying areas around streams):
(1) low deer, low non-endemic insects & pathogens - 4 ha portion of upland forest in the SE quarter of the plot, fenced in YEAR and maintained with only occasional deer presence for the past ## years and with relatively low abundance of canopy species affected by non-endemic pests and pathogens;
(2) high deer, low non-endemic insects & pathogens # ha of upland forest outside the deer exclosure with relatively low abundance of canopy species affected by non-endemic pests and pathogens; 
(3) high deer, high non-endemic insects & pathogens - a # ha section of upland forest outside the deer exclosure with relatively high abundance of canopy species affected by non-endemic pests and pathogens. 
These areas were delineated as follows.
Upland forest was defined according to topographic wetness index, originally calculated for the SCBI plot in @mcgregor_tree_2021.
Vulnerability to non-endemic insects & pathogens was defined based on the abundance of canopy species affected by non-endemic pests and pathogens at the time of plot establishment (2008).

![**Figure `r map_and_photos`. Photos within the SCBI ForestGEO plot: (a) low deer, low non-endemic insects & pathogens, (b) high deer, low non-endemic insects & pathogens, (c) high deer, high non-endemic insects & pathogens ** *(add a map of the plot)* All photos taken September 2023 by K. Anderson-Teixeira.](display/Figure1.png)

@@

A number of field surveys are regularly conducted at the site, two of which are included in our analysis.  As part of the ForestGEO network, the plot undergoes a comprehensive woody plant inventory every five years, according to the protocol outlined in Condit (1998).  This inventory, hereby referred to as the census, includes all stems greater than 1 centimeter in diameter breast height (1.3 meters; dbh).  The census records information regarding the dbh, species, status and location of each stem included in the survey.  Each invidual is assigned an identifying number for sequential data collection in subsequent censuses and outfitted with a metal tag in the field. In multi-stemmed individuals, each additional stem that surpasses the 1 cm dbh threshold receives a stem number and associated tag.  Location is recorded on a digitized map of the plot on the ESRI ArcGIS FieldMaps application (*citation*).  Established in 2008, there have been four censuses at the site, comprising 15 years of detailed forest dynamics data. In addition to the ForestGEO census, since 2014 an annual mortality survey has been conducted on the site. Initiated, in part, as a response to the presence of Emerald Ashborer, the mortality survey is conducted on all trees that had a dbh greater than 10 centimeters in the previous full census. Through assessing individual tree health, this study illustrates trends in forest mortality and identifies factors that are associated with death. Data is collected on the current status, canopy position, percentage of crown living and intact, and visible indicators of poor tree health, such as physical damage, potential pathogens and insect infestation.

@@

* We  
*

@@

For the first time since the establishment of  the SCBI ForestGEO site, whole-plot aboveground C stocks decreased (Figure 2). From 2018 to 2023, whole-plot C stocks have decreased from **"X to X"** Mg C/ha. Both low-nuisance regions show similar decreases in C stocks going from **"X to X"** and **"X to X"** Mg C/ha, respectively. In the high deer high nuisance regions C stocks have decreased for the past two censuses, with a higher magnitude decrease of **"X to X"** Mg C/ha from 2018 to 2023. C stock decrease **is related to/ is explained by** above ground net biomass change. In the first two census intervals, 2008 to 2013 and 2013 to 2018, there were positive C fluxes, but for the 2018 to 2023 interval there was a negative C flux of **VALUE** Mg C/ha/year (Figure 3a).

![**Figure `r biomass_trends`. Aboveground biomass decreased at the site for the first time since establishment. The net decrease at the plot-level was driven by large losses in regions with high vulnerability to nuisance species.**](display/Figure2.jpeg)

The trend in net woody production can be explained by its constituent fluxes (Figure 3). Trends in woody mortality show that for the first two census intervals, the whole plot has lost **VALUE** and then **VALUE** Mg C/ha/year, increasing to a loss of **VALUE** Mg C/ha/year for the most recent census interval (Figure 3b). Woody productivity contributes positively to the flux, showing that the whole plot gained **VALUE** Mg C/ha/year in the first interval, **VALUE** Mg C/ha/year in the second, and **VALUE** Mg C/ha/year in the third (Figure 3c). Areas of low deer, low nuisance species were consistently the most productive ending at **VALUE** Mg C/ha/year in the interval 2018 to 2023, and areas of high deer, high nuisance species were the least productive at **VALUE** Mg C/ha/year in the interval 2018 to 2023. Recruitment has consistently added smaller magnitude amounts of carbon with the whole plot gaining **VALUE** Mg C/ha/year (Figure 3d). Recruitment was highest at **VALUE** Mg C/ha/year in areas with high deer and high vulnerable species, and lowest at **VALUE** Mg C/ha/year in areas with high deer and low vulnerable species.

![**Figure `r loss_gain_net`. Fluxes of aboveground woody biomass. Large increases in woody mortality combined with sustained decreases in woody productivity result in decreasing net biomass in the most recent census.**](display/Figure3.jpeg)

## Species composition of mortality flux

Woody mortality was the primary driver of the negative trend in biomass, and certain species were responsible for higher levels of mortality. Fraxinus americana had the highest aboveground woody mortality rates of any species present on the plot, peaking at **VALUE** Mg C/ha/year lost in *2020* , and **VALUE** Mg C/ha/year lost in 2022 (Figure 4). Quercus velutina contributed the second-highest levels of mortality, losing **VALUE** Mg C/ha/year in 2022. Quercus rubra, Carya species, Quercus prinus, Quercus alba, and Liriodendron tulipifera were all significant contributors as well.

![**Figure `r mortality_by_genera`. Annual woody mortality flux by species/genera. We report large increases in mortality of white ash, red oak, and black oak.**](display/Figure4.jpeg)

## Species composition of recruitment flux
Certain species dominated aboveground woody recruitment, with understory species being recruited at higher rates than canopy species (Figure 5). Out of the canopy species, Carya species showed the highest levels of recruitment with a rate of **VALUE** Mg C/ha/year recruited from 2018 to 2023 (Figure 5a). Acer negundo had the second highest rate of recruitment for canopy species at **VALUE** Mg C/ha/year from 2018 to 2023. Fagus grandifolia, Fraxinus americana, and Ulmus rubra are the other species with the highest recruitment rates. Most canopy species had lower recruitment rates in the 2018 to 2023 census interval than the 2008 to 2013 interval. Lindera benzoin had by far the highest recruitment rate of the understory species at **VALUE** Mg C/ha/year from 2018 to 2023 (Figure 3b). Asimina triloba had the second highest rate of recruitment of the understory species at **VALUE** Mg C/ha/year recruited from 2018 to 2023. Other understory species with high recruitment were Carpinus caroliniana, Hamamelis virginiana, and Celtis occidentalis.

![**Figure `r recruitment_by_genera`. Annual recruitment of canopy & understory species/genera across all censuses. Recruitment of canopy species is significantly higher within the deer exclosure. Recruitment of understory species is 20x that of canopy species.**](display/Figure5.jpeg)

@@

![**Figure `r regen_failure`.**](display/Figure6.jpeg)

@@

# Conclusions /Future Directions
